SU($n$) spin models near the Heisenberg critical point

### 2020-04-01 collective_sun/sun_perturbation.tex
add sketch of multi-body excitation manifolds

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the multi-body operator $\hat\O$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta$ with respect to $H_0$.  These
states are generated by the action of an $M$-body operator
$X_{\m w_\Delta}$ on a PS state $\ket\psi\in\M_0$.  We denote the
manifold of all states that can be generated by the action of some
$M$-body operator, but not any $r$-body operator with $r<M$, by
$\M_M$; we will refer to $\M_M$ as the $M$-body excitation manifold.
We then denote the a direct sum of all $\M_r$ for $r\le M$ by
$\ul\M_M$, i.e.~$\ul\M_M\equiv\bigoplus_{r\le M}\M_r$.  The manifold
$\ul\M_M$ consists of all states accessible by the action of $r$-body
operators with $r\le M$ on PS states in $\M_0$.

By construction, the $M$-body operator $X_{\m w}$ generally couples PS
states in $\M_0$ to asymmetric states in $\ul\M_M\setminus\M_0$.  If
the coupling between $\M_0$ and its complement is weak compared to the
spectral gap $\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect
of $X_{\m w}$ perturbatively and derive an effective Hamiltonian
$H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory
formally breaks down when the operator norm
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
$X_{\m w}$ first couples $\M_0$ to $\ul\M_M$, then $\ul\M_M$ to
$\ul\M_{2M}$, and so on.  Even if
$\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
incurred by $H_0$ from breaking the permutational symmetry of an
initial state $\ket\psi\in\M_0$ suppresses the leakage of population
into manifolds $\M_r$ of increasing $r$.  One might therefore capture
the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
$X_{\m w}$ onto $\ul\M_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
resulting in a theory that resembles first-order perturbation theory
on $\ul\M_{r_{\t{max}}}$.  In the perturbative regime
$\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in a formal
perturbation theory on $\M_0$.  In the non-perturbative regime
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, the error incurred by
truncating the Hilbert space of a spin system past
$\ul\M_{r_{\t{max}}}$ is not clear a priori.  Nonetheless, simulations
restricted to $\ul\M_{r_{\t{max}}}$ can be benchmarked by examining
the maximum populations within all $\M_r$ for $r\le r_{\t{max}}$
throughout simulation.  As long as
\begin{enumerate*}
\item these populations fall off with increasing $r$,
\item and the maximal population within large-$r$ manifolds $\M_r$ is
  small,
\end{enumerate*}
one should expect such simulations to faithfully capture (at least
qualitatively) the dynamical behavior of the spin system under
consideration.

>>>>

% for energy diagrams, etc.
\usepackage{tikz}
\usetikzlibrary{arrows,math}
\tikzstyle{every picture}+=[remember picture]

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the multi-body operator $\hat\O$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta$ with respect to $H_0$.  These
states are generated by the action of an $M$-body operator
$X_{\m w_\Delta}$ on a PS state $\ket\psi\in\M_0$.  We denote the
manifold of all states that can be generated by the action of some
$M$-body operator, but not any $r$-body operator with $r<M$, by
$\M_M$; we will refer to $\M_M$ as the $M$-body excitation manifold.
We then denote the a direct sum of all $\M_r$ for $r\le M$ by
$\ul\M_M$, i.e.~$\ul\M_M\equiv\bigoplus_{r\le M}\M_r$.  The manifold
$\ul\M_M$ consists of all states accessible by the action of $r$-body
operators with $r\le M$ on PS states in $\M_0$.  These definitions are
sketched out in Figure \ref{fig:manifold_sketch}.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \draw[ultra thick] (0,0) -- node[above] {$\M_0$} ++(10em,0);
    \draw[ultra thick] (0,3em) -- node[above] {$\M_1$} ++(10em,0);
    \draw[ultra thick] (0,2*3em) -- node[above] {$\M_2$} ++(10em,0);
    \draw[->] (-.2em,0) to [bend left] (-.2em,3em);
    \node (X) at (.2em,0.5*3em) {$\hat\O_1$};
    \draw[->] (-.4em,0) to [bend left] (-.2em,2*3em);
    \node (XY) at (-2em,3em) {$\hat\O_2$};
    \draw (10em+.2em,-.2em) -- (10em+.4em,-.2em)
    -- (10em+.4em,3em+.2em) -- (10em+.2em,3em+.2em);
    \node (cM1) at (10em+1.5em,0.5*3em) {$\col\M_1$};
    \draw (10em+2.4em+.2em,-.4em) -- (10em+2.4em+.4em,-.4em)
    -- (10em+2.4em+.4em,6em+.4em) -- (10em+2.4em+.2em,6em+.4em);
    \node (cM1) at (10em+2.4em+1.5em,3em) {$\col\M_2$};
  \end{tikzpicture}
  \caption{Schematic of low-lying multi-body manifolds.  Starting with
    the permutationally-state manifold $\M_0$, the single-body
    excitation manifold $\M_1$ consists of states accessible via
    single-body perturbations $\hat\O_1$.  The two-body excitation
    manifold $\hat\O_2$ then consists of states accessible via
    two-body perturbations, but not single-body perturbations.  The
    joint manifold $\col\M_M$ consists of all states accessible by
    operators $r$-body operators $\O_r$ with $r\le M$.}
  \label{fig:manifold_sketch}
\end{figure}

By construction, the $M$-body operator $X_{\m w}$ generally couples PS
states in $\M_0$ to asymmetric states in $\ul\M_M$.  If the coupling
between $\M_0$ and its complement is weak compared to the spectral gap
$\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect of
$X_{\m w}$ perturbatively and derive an effective Hamiltonian
$H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory
formally breaks down when the operator norm
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
$X_{\m w}$ first couples $\M_0$ to $\ul\M_M$, then $\ul\M_M$ to
$\ul\M_{2M}$, and so on.  Even if
$\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
incurred by $H_0$ from breaking the permutational symmetry of an
initial state $\ket\psi\in\M_0$ suppresses the leakage of population
into manifolds $\M_r$ of increasing $r$.  One might therefore capture
the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
$X_{\m w}$ onto $\ul\M_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
resulting in a theory that resembles first-order perturbation theory
on $\ul\M_{r_{\t{max}}}$.  In the perturbative regime
$\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in a formal
perturbation theory on $\M_0$.  In the non-perturbative regime
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, the error incurred by
truncating the Hilbert space of a spin system past
$\ul\M_{r_{\t{max}}}$ is not clear a priori.  Nonetheless, simulations
restricted to $\ul\M_{r_{\t{max}}}$ can be benchmarked by examining
the maximum populations within all $\M_r$ for $r\le r_{\t{max}}$
throughout simulation.  As long as
\begin{enumerate*}
\item these populations fall off with increasing $r$,
\item and the maximal population within large-$r$ manifolds $\M_r$ is
  small,
\end{enumerate*}
one should expect such simulations to faithfully capture (at least
qualitatively) the dynamical behavior of the spin system under
consideration.
